manuscript_figures

### 2024-06-19 scripts/manuscript_figures.Rmd
made temp analysis and table

# Load libraries 
```{r}
library(plyr)
library(dplyr)
library(tigris)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Read the fish aggregated CSV file 2019-2023 and skip the first column
fish <- read.csv("data/kelp_timeseries_data/fish_Aggregated.csv") %>%
          select(-1)

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_cover + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover), y = log(mean_fish_biomass))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass) ~ mean_kelp_biomass + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass), y = log(mean_fish_biomass))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

@@

# Create scatter plot
figure.4 <- ggplot(aggregated.data.rug, aes(x = mean_rugosity, y = mean_fish_biomass)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_rugosity, 2)), 
           hjust = 2.1, vjust = 1.1, size = 4, parse = FALSE) +
  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Rugosity Complexity") +
  theme_classic();figure.4

@@

# output to .csv
write.csv("output/cleaned_data/mast_output_dataframe.csv")

@@

ggsave(filename = "output/figures/correlation_analysis_figure.png", device = "png", width = 10, height = 10)

@@

# Figure 5: Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
# combine all dataframes
```{r}
# Rename columns to remove spaces and special characters
names(all.params.data) <- c("YEAR", "SITE", "Fish_Biomass", "Rugosity", "Kelp_Cover", "Kelp_Biomass")

@@

# Fit a multiple regression model
fit1 <- lm(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover), data = all.params.data)

@@

# run with just site as random effect
# Fit the mixed-effects model
fit4 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | SITE), data = all.params.data)

@@

# run with just year as random effect
# Fit the mixed-effects model
fit5 <- lmer(log(Fish_Biomass) ~ log(Rugosity) * log(Kelp_Biomass) * log(Kelp_Cover) + 
              (1 | YEAR), data = all.params.data)

@@

# Run normality on the temp data
qqPlot(temperature_data$Temp, main = "QQ Plot of Temperature Data")

# Data is not normal so log transform
qqPlot(temperature_data$log(Temp), main = "QQ Plot of Temperature Data")

# Fit linear model
lm_model <- lm(log(Temp) ~ Date + Site, data = temperature_data)

# Summary of the model
summary(lm_model)

>>>>

# Load libraries 
```{r}
library(plyr)
library(kableExtra)
library(gt)
library(dplyr)
library(tigris)
library(MASS)
library(htmlwidgets)
library(car)
library(ggiraphExtra)
library(MuMIn)
library(lmerTest)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(ggmap)
library(lubridate)
library(cowplot)
library(ggplot2)
library(PerformanceAnalytics)
library(reshape2)
library(viridis)
library(GGally)
library(corrplot)
library(patchwork)
library(broom)
library(knitr)
library(ggspatial)
library(rslp)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)

@@

# Read the fish aggregated CSV file 2019-2023 and skip the first column
fish <- read.csv("data/kelp_timeseries_data/fish_Aggregated.csv")

# remove Xcolumn, not needed
fish <- fish[ -c(1) ]

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_cover + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated_data_cover)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Create scatter plot
figure.2 <- ggplot(aggregated_data_cover, aes(x = log(mean_kelp_cover + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_cover, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp cover (%)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Cover") +
  theme_classic();figure.2

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_kelp_biomass + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)
#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Create scatter plot
figure.3 <- ggplot(aggregated.data.morph, aes(x = log(mean_kelp_biomass + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_kelp_biomass, 3)), hjust = 1.4, vjust = 1.1, size = 4, parse = FALSE) + # Adjust vjust
  labs(x = "Log Kelp Biomass (g/m2)", y = "Log Fish biomass (g/m2)",
       title = "Fish Biomass vs Kelp Biomass") +
  theme_classic();figure.3

@@

# Fit the linear mixed effects model with log transformation
lmm_model <- lmer(log(mean_fish_biomass + 1.1) ~ log(mean_rugosity + 1.1) + SITE + (1 | SITE/YEAR), data = aggregated.data.morph)

#check normality
qqPlot(residuals(lmm_model)) #improved to normal

@@

# Create scatter plot
figure.4 <- ggplot(aggregated.data.rug, aes(x = log(mean_rugosity + 1.1), y = log(mean_fish_biomass + 1.1))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, formula = y ~ x, color = "blue") +
  annotate("text", x = Inf, y = Inf, label = paste0("Marginal R^2 = ", marginal_r2, "\nConditional R^2 = ", conditional_r2, "\np = ", signif(p_value_rugosity, 2)), 
           hjust = 2.1, vjust = 1.1, size = 4, parse = FALSE) +
  labs(x = "Rugosity Complexity", y = "Fish biomass (g/m2)",
       title = "Fish Biomass vs Rugosity Complexity") +
  theme_classic();figure.4

@@

# output to .csv
write.csv(master.dat, "output/cleaned_data/master_output_dataframe.csv")

@@

ggsave(filename = "output/figures/figure_5.png", device = "png", width = 10, height = 10)

@@

# Multiple regression between fish biomass (g/m2), kelp biomass (g/m2), kelp cover, rugosity, temperature in ggplot 
# combine all dataframes
```{r}
# Rename columns to remove spaces and special characters
names(all.params.data) <- c("YEAR", "SITE", "Fish_Biomass", "Rugosity", "Kelp_Cover", "Kelp_Biomass")

@@

# Fit a multiple regression model
fit1 <- lm(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1), data = all.params.data)

@@

# run with just site as random effect
# Fit the mixed-effects model
fit4 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
              (1 | SITE), data = all.params.data)

@@

# run with just year as random effect
# Fit the mixed-effects model
fit5 <- lmer(log(Fish_Biomass + 1.1) ~ log(Rugosity + 1.1) * log(Kelp_Biomass + 1.1) * log(Kelp_Cover + 1.1) + 
              (1 | YEAR), data = all.params.data)

@@

# Check normality for each site
shapiro.test(temperature_data$Temp[temperature_data$Site == "KB"])
shapiro.test(temperature_data$Temp[temperature_data$Site == "FW"])

# data not normal

# Perform Mann-Whitney U Test for site
mann_whitney_test <- wilcox.test(Temp ~ Site, data = temperature_data, exact = FALSE)
print(mann_whitney_test)

# Convert Date column to Date class
temperature_data$Date <- as.Date(temperature_data$Date, format = "%m/%d/%y")

# Extract Year and Month
temperature_data <- temperature_data %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  )

# Calculate monthly averages
monthly_averages <- temperature_data %>%
  group_by(Year, Month, Site) %>%
  summarize(AverageTemp = mean(Temp), .groups = 'drop')

# Fit linear model
model <- lm(log(AverageTemp +1.1) ~ Month + Year + Site, data = monthly_averages)

#check normality
qqPlot(residuals(model)) #improved to normal

#not normal

# Remove specific outliers
cleaned_data <- monthly_averages[-c(95, 96), ]

# Refit the model without outliers
cleaned_model <- lm(log(AverageTemp + 1.1) ~ Month + Year + Site, data = cleaned_data)
qqPlot(residuals(cleaned_model))
summary(cleaned_model)

#month is sig different but year and site are not

# make a table for temperature values (monthly mean plus/minus one standard error) for each site (for supplement)

# Calculate monthly mean and standard error
summary_table <- temperature_data %>%
  group_by(Site, Month) %>%
  summarize(
    MeanTemp = mean(Temp),
    StdDev = sd(Temp),
    N = n(),
    StdError = StdDev / sqrt(N)
  ) %>%
  mutate(
    MeanPlusSE = MeanTemp + StdError,
    MeanMinusSE = MeanTemp - StdError
  ) %>%
  ungroup() %>%
  dplyr::select(Site, Month, MeanTemp, MeanPlusSE, MeanMinusSE)

# Mapping month numbers to month names
month_names <- c("January", "February", "March", "April", "May", "June", 
                 "July", "August", "September", "October", "November", "December")

# Mapping site abbreviations to full names
site_names <- c("FW" = "Fort Wetherill", "KB" = "King's Beach")

# Replace numeric month values with month names
summary_table$Month <- month_names[summary_table$Month]

# Replace site abbreviations with full names
summary_table$Site <- site_names[summary_table$Site]

# Create a gt table
table_gt <- gt(summary_table) %>%
  tab_header(
    title = "Monthly Mean Temperatures by Site",
    subtitle = "Summary Table"
  ) %>%
  fmt_number(
    columns = vars(MeanTemp, MeanPlusSE, MeanMinusSE),
    decimals = 1
  ) %>%
  tab_style(
    style = list(
      cell_text(color = "black", weight = "bold")
    ),
    locations = cells_body()
  );table_gt

# Save as HTML
gtsave(table_gt, file = "output/temperature_table.html")

```
Temperature between sites is not significantly different

	Wilcoxon rank sum test with continuity correction

data:  Temp by Site
W = 4273000, p-value = 0.9511
alternative hypothesis: true location shift is not equal to 0
